---
title: "Data Analytics Portfolio"
author: "Anthony M. Mugambi"
date: "2026-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##  Profile

Data analyst with strong expertise in statistical analysis, econometrics, and data-driven research. Proficient in R programming, data visualization, and regression modeling, with hands-on experience analyzing real-world datasets. Demonstrated ability to transform data into actionable insights for research and policy applications.

## ️ Technical Skills

- R Programming (ggplot2, dplyr, readxl, MASS)
- Data Cleaning and Transformation
- Statistical Analysis (Chi-square, Fisher’s Exact Test)
- Regression Modeling (Logistic & Ordered Logistic)
- Data Visualization
- Research & Academic Writing 

##  Data Loading

```{r}

library(readxl)

```{r echo=FALSE, results='hide'}
apiculture_data <- read_excel("C:/Users/pc/Documents/2025/data apiculture/apiculturemay2026.xlsx")

# confirm it loaded
head(apiculture_data)
```

```{r}

table(apiculture_data$ANNUALINCOME)
```
```

##  Project: Apiculture Income Analysis (Meru County)

This project analyzes the determinants of income among small-scale beekeepers using survey data. The analysis focuses on the role of training, technology adoption, and education in influencing income outcomes.

### Methods Used:
- Descriptive statistics
- Chi-square and Fisher’s Exact Test
- Ordered logistic regression



##  Sample Analysis


## Including Plots
##  Data Visualization

```{r}
library(ggplot2)

ggplot(apiculture_data, aes(x = ANNUALINCOME)) +
  geom_bar() +
  theme_minimal() +
  labs(
    title = "Income Distribution of Beekeepers",
    x = "Income Category",
    y = "Frequency"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Regression Model

```{r}
apiculture_data$ANNUALINCOME <- factor(
  apiculture_data$ANNUALINCOME,
  levels = c("<5000", "5000-10000", "11000-20000", "21000-100000", "100001+"),
  ordered = TRUE
)
```


```{r echo=FALSE, results='hide'}
library(MASS)

model <- polr(
  ANNUALINCOME ~ TYPEOFHIVES + FORMALBEEKEEPINGTRAINING,
  data = apiculture_data
)

summary(model)
```


##  Key Findings

- Majority of beekeepers fall in low-income category
- Training significantly improves income outcomes
- Traditional hives reduce productivity
- Education has a weaker effect compared to training

## Research Output

- Determinants of Income in Apiculture 
- Ongoing: Adoption of Modern Beekeeping Practices

## Contact

- Email: antoinemuriungi@gmail.com  
- Institution: Meru University of Science and Technology  


